Expression QTL hotspot dissection with mediation analysis and pleiotropy testing

### 2019-09-22 analysis/paper/Rmd/hotspot-one-chromosome.Rmd
removed, partially, code dealing with allele effects.

```{r, prepare-annotations-{{hot_chr}}}
# need 5 annotation object: 1. local v nonlocal, 2. mediation LOD difference for all local traits, 3. mediation LOD difference proportion for all local traits, 4. 8 allele effects per trait, 5. peak position per trait, encoded as binned position
# 2,3,4 are numeric; 1 & 5 are binary
## local_nonlocal_annot already has the allele effect and local v nonlocal; binning has the peak position. Need to join these two on symbol
### med_tib has lod diff & lod diff proportion for all local - nonlocal pairs
fake_tibble_mediation <- tidyr::crossing(symbol = local_annot2$symbol, 
                                         symbol2 = local_annot2$symbol) %>%
  mutate(dat = NA) %>%
  spread(key = symbol, value = dat) %>%
  rename(symbol = symbol2)
# to add to mediation tibbles

# first med tibble
lod_diffs_tibble <- med_tib %>%
  select( - c(lod_without_med, 
              lod_with_med, 
              lod_diff_proportion, 
              local_gene_id, 
              nonlocal_gene_id)
          ) %>%
  tidyr::spread(key = local_symbol, 
                value = lod_diff) %>% 
  dplyr::rename(symbol = nonlocal_symbol) %>%
  dplyr::bind_rows(fake_tibble_mediation)
# second med tibble
lod_diff_proportions_tibble <- med_tib %>%
  select( - c(lod_without_med, 
              lod_with_med, 
              lod_diff, 
              local_gene_id, 
              nonlocal_gene_id)
          ) %>%
  tidyr::spread(key = local_symbol, 
                value = lod_diff_proportion) %>% # 147 rows by 32 columns
  rename(symbol = nonlocal_symbol) %>%
  bind_rows(fake_tibble_mediation)
# effects tibble
allele_effects_tibble <- local_nonlocal_annot %>%
  dplyr::select(symbol, 18:25)
# peak positions tibble
peak_positions <- binning %>%
  select(symbol, pos)
# local indicator
local_tibble <- local_nonlocal_annot %>%
  select(symbol, local_tx) %>% 
  ungroup()
# univariate lods
lod_tibble <- local_nonlocal_annot %>%
  select(symbol, lod) %>%
  ungroup()
```

```{r, define-contrasts-{{hot_chr}}}
contrasts_BC_FG <- local_nonlocal_annot %>%
  ungroup() %>%
  select(symbol, cA:cH, local_tx, lod) %>%
  mutate(row_num = 1:nrow(local_nonlocal_annot), 
         AB = abs(cA - cB),
         AC = abs(cA - cC),
         AD = abs(cA - cD),
         AE = abs(cA - cE),
         AF = abs(cA - cF),
         AG = abs(cA - cG),
         AH = abs(cA - cH),
         BC = abs(cB - cC),
         BD = abs(cB - cD),
         BE = abs(cB - cE),
         BF = abs(cB - cF),
         BG = abs(cB - cG),
         BH = abs(cB - cH),
         CD = abs(cC - cD),
         CE = abs(cC - cE),
         CF = abs(cC - cF),
         CG = abs(cC - cG),
         CH = abs(cC - cH),
         DE = abs(cD - cE),
         DF = abs(cD - cF),
         DG = abs(cD - cG),
         DH = abs(cD - cH),
         EF = abs(cE - cF),
         EG = abs(cE - cG),
         EH = abs(cE - cH),
         FG = abs(cF - cG),
         FH = abs(cF - cH),
         GH = abs(cG - cH)
  ) %>%
  mutate(BC_A = abs((cB + cC) / 2 - cA),
         BC_D = abs((cB + cC) / 2 - cD),
         BC_E = abs((cB + cC) / 2 - cE),
         BC_F = abs((cB + cC) / 2 - cF),
         BC_G = abs((cB + cC) / 2 - cG),
         BC_H = abs((cB + cC) / 2 - cH),
         BC_FG = abs((cB + cC) - (cF + cG))
         ) %>%
  select(symbol, BC_FG)
```

```{r, order-annotations-{{hot_chr}}}
# all annots 
annots_ord <- list(lod_diffs_tibble, 
                   lod_diff_proportions_tibble, 
                   allele_effects_tibble,
                   peak_positions, 
                   local_tibble,
                   lod_tibble, 
                   contrasts_BC_FG
                    ) %>% 
  purrr::map(.f = function(df, rn){
    out <- df[match(x = rn, table = df$symbol), ]
    return(out)
  }, 
  rn = rn)
```

@@

```{r iheatmapr-{{hot_chr}}, out.width="25in", out.height = "10in"}
ann3 <- annots_ord[[3]] %>%
  ungroup() %>%
  select(- symbol) %>%
  as.matrix() %>%
  (function(x){rownames(x) <- annots_ord[[3]]$symbol; return(x)})
colnames(ann3) <- LETTERS[1:8]
ann2 <- annots_ord[[2]] %>%
  select(-symbol) %>%
  as.matrix() %>%
  (function(x){rownames(x) <- annots_ord[[2]]$symbol; return(x)})
ann1 <- annots_ord[[1]] %>%
  select(-symbol) %>%
  as.matrix() %>%
  (function(x){rownames(x) <- annots_ord[[1]]$symbol; return(x)})
ann5 <- annots_ord[[5]] %>%
  select(-symbol) %>%
  as.matrix() %>%
  (function(x){rownames(x) <- annots_ord[[5]]$symbol; return(x)})
ann6 <- annots_ord[[6]] %>%
  select(-symbol) %>%
  as.matrix() %>%
  (function(x){rownames(x) <- annots_ord[[6]]$symbol; return(x)})
ann7 <- annots_ord[[7]] %>% # BC_FG contrast
  select(-symbol) %>%
  as.matrix() %>%
  (function(x){rownames(x) <- annots_ord[[7]]$symbol; return(x)})
ann4 <- annots_ord[[4]] %>%
  ungroup() %>%
  select(-symbol) %>%
  as.matrix() %>%
  (function(x){rownames(x) <- annots_ord[[4]]$symbol; return(x)})
mymat <- pm[ord, ord] %>% # order both cols and rows by 'ord'
  apply(FUN = trunc2, MARGIN = 2, threshold = 5)
iheatmapr::main_heatmap(mymat, show_colorbar = FALSE, layout = list(font = list(size = 4))) %>%
  iheatmapr::add_col_labels() %>%
  iheatmapr::add_col_annotation(ann5) %>%
  iheatmapr::add_main_heatmap(ann3, name = "signal3", size = 0.2, show_colorbar = FALSE) %>%
    iheatmapr::add_col_labels() %>%

  iheatmapr::add_main_heatmap(ann2, name = "signal2", size = 0.5, show_colorbar = FALSE) %>%
  iheatmapr::add_col_labels() %>%
  iheatmapr::add_main_heatmap(ann1, name = "signal1", size = 0.5, show_colorbar = FALSE) %>%
  iheatmapr::add_row_annotation(ann5, side = "left") %>%
    iheatmapr::add_col_labels() %>%
  iheatmapr::add_row_annotation(ann4, side = "left") %>%
  iheatmapr::add_row_annotation(ann6, side = "left") %>%
  iheatmapr::add_row_annotation(ann7, side = "left") %>%
  iheatmapr::add_row_labels()

```

>>>>

```{r, prepare-annotations-{{hot_chr}}}
# need 5 annotation object: 1. local v nonlocal, 2. mediation LOD difference for all local traits, 3. mediation LOD difference proportion for all local traits, 4. 8 allele effects per trait, 5. peak position per trait, as a one-column heatmap
# 2,3,4 are numeric; 1 & 5 are binary
## local_nonlocal_annot already has the allele effect and local v nonlocal; binning has the peak position. Need to join these two on symbol
### med_tib has lod diff & lod diff proportion for all local - nonlocal pairs
fake_tibble_mediation <- tidyr::crossing(symbol = local_annot2$symbol, 
                                         symbol2 = local_annot2$symbol) %>%
  mutate(dat = NA) %>%
  spread(key = symbol, value = dat) %>%
  rename(symbol = symbol2)
# to add to mediation tibbles

# first med tibble
lod_diffs_tibble <- med_tib %>%
  select( - c(lod_without_med, 
              lod_with_med, 
              lod_diff_proportion, 
              local_gene_id, 
              nonlocal_gene_id)
          ) %>%
  tidyr::spread(key = local_symbol, 
                value = lod_diff) %>% 
  dplyr::rename(symbol = nonlocal_symbol) %>%
  dplyr::bind_rows(fake_tibble_mediation)
# second med tibble
lod_diff_proportions_tibble <- med_tib %>%
  select( - c(lod_without_med, 
              lod_with_med, 
              lod_diff, 
              local_gene_id, 
              nonlocal_gene_id)
          ) %>%
  tidyr::spread(key = local_symbol, 
                value = lod_diff_proportion) %>% # 147 rows by 32 columns
  rename(symbol = nonlocal_symbol) %>%
  bind_rows(fake_tibble_mediation)
# effects tibble
allele_effects_tibble <- local_nonlocal_annot %>%
  dplyr::select(symbol, 18:25)
# peak positions tibble
peak_positions <- local_nonlocal_annot %>%
  select(symbol, pos)
# local indicator
local_tibble <- local_nonlocal_annot %>%
  select(symbol, local_tx) %>% 
  ungroup()
# univariate lods
lod_tibble <- local_nonlocal_annot %>%
  select(symbol, lod) %>%
  ungroup()
```

```{r, order-annotations-{{hot_chr}}}
# all annots 
annots_ord <- list(lod_diffs_tibble, 
                   lod_diff_proportions_tibble, 
                   allele_effects_tibble,
                   peak_positions, 
                   local_tibble,
                   lod_tibble 
                    ) %>% 
  purrr::map(.f = function(df, rn){
    out <- df[match(x = rn, table = df$symbol), ]
    return(out)
  }, 
  rn = rn)
```

@@

```{r {{hot_chr}}-ann3}
ann3 <- annots_ord[[3]] %>%
  ungroup() %>%
  select(- symbol) %>%
  as.matrix() %>%
  (function(x){rownames(x) <- annots_ord[[3]]$symbol; return(x)})
colnames(ann3) <- LETTERS[1:8]
```

```{r {{hot_chr}}-ann2}
ann2 <- annots_ord[[2]] %>%
  select(-symbol) %>%
  as.matrix() %>%
  (function(x){rownames(x) <- annots_ord[[2]]$symbol; return(x)})
```

```{r {{hot_chr}}-ann1}
ann1 <- annots_ord[[1]] %>%
  select(-symbol) %>%
  as.matrix() %>%
  (function(x){rownames(x) <- annots_ord[[1]]$symbol; return(x)})
```

```{r {{hot_chr}}-ann5}
ann5 <- annots_ord[[5]] %>%
  select(-symbol) %>%
  as.matrix() %>%
  (function(x){rownames(x) <- annots_ord[[5]]$symbol; return(x)})
```

```{r {{hot_chr}}-ann6}
ann6 <- annots_ord[[6]] %>%
  select(-symbol) %>%
  as.matrix() %>%
  (function(x){rownames(x) <- annots_ord[[6]]$symbol; return(x)})
```

```{r {{hot_chr}}-ann4}
ann4 <- annots_ord[[4]] %>%
  ungroup() %>%
  select(-symbol) %>%
  as.matrix() %>%
  (function(x){rownames(x) <- annots_ord[[4]]$symbol; return(x)})
```

```{r iheatmapr-{{hot_chr}}, out.width="25in", out.height = "10in"}
mymat <- pm[ord, ord] %>% # order both cols and rows by 'ord'
  apply(FUN = trunc2, MARGIN = 2, threshold = 5)
iheatmapr::main_heatmap(mymat, show_colorbar = FALSE, layout = list(font = list(size = 4))) %>%
  iheatmapr::add_col_labels() %>%
  iheatmapr::add_col_annotation(ann5) %>%
  iheatmapr::add_main_heatmap(ann3, name = "signal3", size = 0.2, show_colorbar = FALSE) %>%
    iheatmapr::add_col_labels() %>%

  iheatmapr::add_main_heatmap(ann2, name = "signal2", size = 0.5, show_colorbar = FALSE) %>%
  iheatmapr::add_col_labels() %>%
  iheatmapr::add_main_heatmap(ann1, name = "signal1", size = 0.5, show_colorbar = FALSE) %>%
  iheatmapr::add_row_annotation(ann5, side = "left") %>%
    iheatmapr::add_col_labels() %>%
  iheatmapr::add_row_annotation(ann4, side = "left") %>%
  iheatmapr::add_row_annotation(ann6, side = "left") %>%
  iheatmapr::add_row_labels()
```
